---
title: "Replication and Extension"
author: "Sheridamae Gudez"
date: "`r Sys.Date()`"
output: html_document
---
*Replication Article*
Grund, T. and Densley, J. (2015) Ethnic Homophily and Triad Closure: Mapping Internal Gang Structure Using Exponential Random Graph Models. Journal of Contemporary Criminal Justice, Vol. 31, Issue 3, pp. 354-370
"Previous studies indicate the importance of similarities between street gang members in terms of ethnicity for mapping the patterns of co-offending relationships within gangs. Individual members are more likely to co-offend with other members who are from the same ethnicity. Past studies, however, do not appropriately account for the fact that correlation between attributes of co-offending gang members may be driven by alternative mechanisms. Most importantly, the presence of clustering in networks can dramatically affect the assessment and make us believe that homophily—the deliberate choice to co-offend with others from the same ethnic group—is important while in fact it is not. In this article, we recreate the internal structure of a London-based street gang with exponential random graph (ERG) models. Our results refine the role of ethnicity for co-offending within gangs. While homophily is still prevalent, the effect diminishes when triad closure—the tendency for two individuals to offend with each other when they also offend with a common third person—is considered. Furthermore, we extend existing ERG specifications and investigate the interaction between ethnic homophily and triad closure. Findings indicate that ethnic homophily is even stronger when it is embedded in co-offending triads"

*Set-Up*
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Install Packages
```{r message=FALSE, warning=FALSE, include=FALSE}
rm(list=ls()) #Clearing environment
libraries <- c("tidyverse","shinyjs", "sna", "intergraph", 
               "GGally", "ggplot2", "tibble", "dplyr", "ggplot2", "texreg",
               "ggpubr", "btergm", "igraph") #check for installed packages and read in libraries
for (i in libraries){
  tryCatch(library(i, character.only = TRUE),
           print(paste(i, "installed")),
           error = function(e) {install.packages(i);
             library(i, character.only = TRUE)}
  )
  print(paste(i, "loaded"))
}
set.seed(5) #set seed for replication
```
Read-In Data
Data is on co-offending in a London-based inner-city street gang, 2005-2009, operating from a social housing estate. Data comes from anonymised police arrest and conviction data for all confirmed members of the gang. 
```{r}
adj <- read.csv("LONDON_GANG.csv") #Edge List
att <- read.csv("LONDON_GANG_ATTR.csv") #Attributes Data
```

*Data Cleaning*
Mutate
```{r}
#Mutating att data#
att$X <- paste0("X", att$X) #Create row names by prepending "X" to the values in the 'X' column
#Mutating adj. matrix nodes#
adj$X <- paste0("X", adj$X) #Create row names by prepending "X" to the values in the 'X' column
adj <- adj %>%
  column_to_rownames(var = "X")
adj <- as.matrix(adj)
```
Creating Relationship Lists
```{r}
#Reshape the matrix into an undirected, binary edge list
relationship <- as.data.frame(adj) %>%
  rownames_to_column(var = "from") %>%
  gather(key = "to", value = "relationship", -from)
relationship <- relationship %>%#Convert relationship values to binary ties and filter
  filter(relationship > 1) %>% # Keep only co-offending relationships
  mutate(pair = pmap_chr(list(from, to), ~ paste(sort(c(...)), collapse = "-"))) %>% # Create a pair identifier while retaining the relationship value
  group_by(pair) %>% # Keep only distinct pairs, but keep the first occurrence of the relationship
  summarise(from = first(from), to = first(to), relationship = first(relationship), .groups = 'drop') %>%
  select(-pair) # Remove the pair column
```

*Create iGraph Objects*
1-Mode matrix 54 x 54 persons by persons, undirected, valued. 
Network tie values:1 (hang out together); 2 (co-offend together); 3 (co-offend together, serious crime); 4 (co-offend together, serious crime, kin).
Attributes: Age, Birthplace (1 = West Africa, 2= Caribbean, 3= UK, 4= East Africa), Residence, Arrests, Convictions, Prison, Music.

Igraph Network
```{r}
igang <- graph_from_adjacency_matrix(adj, mode="undirected", weighted=T) #Create an igraph object
# Apply a threshold to keep only nodes with weight > 1
threshold <- 2
igang <- delete_edges(igang, E(igang)[weight < threshold]) 
igang <- delete_vertices(igang, V(igang)[degree(igang) == 0])
```
Adding Attributes Data
```{r}
#Subset the attributes to match the number of vertices#
# Step 1: Extract the node names from the network object
network_nodes <- as.character(V(igang)$name)  # Ensure it's a character vector
# Step 2: Ensure 'id' in the dataframe is also a character vector
att$X <- as.character(att$X)
# Step 3: Filter the attribute dataframe to only keep rows with nodes present in the network
att <- att %>%
  filter(X %in% network_nodes)
#Edge Attributes#
E(igang)$relationship <- relationship$relationship #apply edge attributes
# Define colors for each type of relationship
edge_colors <- c("green","blue", "yellow", "red") #1 (hang out together); 2 (co-offend together); 3 (co-offend together, serious crime); 4 (co-offend together, serious crime, kin).
#Vertex Attributes#
#Applying attributes#
V(igang)$age <- att$Age # Apply 'age' attribute
V(igang)$residence <- att$Residence # Apply 'residence' attribute
V(igang)$residence <- as.factor(V(igang)$residence)
V(igang)$arrests <- att$Arrests # Apply 'arrests' attribute
V(igang)$convictions <- att$Convictions # Apply 'convictions' attribute
V(igang)$prison <- att$Prison # Apply 'prison' attribute
V(igang)$prison <- as.factor(V(igang)$prison)
V(igang)$music <- att$Music # Apply 'music' attribute
V(igang)$music <- as.factor(V(igang)$music)
V(igang)$rank <- att$Ranking # Apply 'ranking' attribute
# Map Birthplace to shapes
V(igang)$birthplace <- as.factor(att$Birthplace) # Apply 'birthplace' attribute
```
Extract Attribute Data for Descriptives
```{r}
#Extract node attributes
vertex_data <- data.frame(id = V(igang)$name, 
                          age = V(igang)$age, 
                          residence = V(igang)$residence, 
                          arrests = V(igang)$arrests, 
                          convictions = V(igang)$convictions, 
                          prison = V(igang)$prison, 
                          music = V(igang)$music, 
                          rank = V(igang)$rank, 
                          birthplace = V(igang)$birthplace
                          )
```

*Descriptives*
Creating Plot Aesthetics
```{r}
#Creating Birthplace legend levels
birthplace_shapes <- c("circle", "square", "triangle", "diamond")
names(birthplace_shapes) <- c("West Africa", "Caribbean", "UK", "East Africa")
# Map Birthplace to colors
birthplace_colors <- c("orange", "brown", "darkgrey", "purple")
names(birthplace_colors) <- c("West Africa", "Caribbean", "UK", "East Africa")
#Assign vertex shapes and colors based on Birthplace
V(igang)$bshape <- birthplace_shapes[V(igang)$birthplace]
V(igang)$bcolor <- birthplace_colors[V(igang)$birthplace]

# triangle vertex shape
mytriangle <- function(coords, v = NULL, params) {
  vertex.color <- params("vertex", "color")
  if (length(vertex.color) != 1 && !is.null(v)) {
    vertex.color <- vertex.color[v]
  }
  vertex.size <- 1 / 200 * params("vertex", "size")
  if (length(vertex.size) != 1 && !is.null(v)) {
    vertex.size <- vertex.size[v]
  }
  symbols(
    x = coords[, 1], y = coords[, 2], bg = vertex.color,
    stars = cbind(vertex.size, vertex.size, vertex.size),
    add = TRUE, inches = FALSE
  )
}
# clips as a circle
add_shape("triangle",
  clip = shapes("circle")$clip,
  plot = mytriangle
)

mystar <- function(coords, v = NULL, params) {
  vertex.color <- params("vertex", "color")
  if (length(vertex.color) != 1 && !is.null(v)) {
    vertex.color <- vertex.color[v]
  }
  vertex.size <- 1 / 200 * params("vertex", "size")
  if (length(vertex.size) != 1 && !is.null(v)) {
    vertex.size <- vertex.size[v]
  }
  
# Draw a diamond (rotated square)
mapply(coords[, 1], coords[, 2], vertex.color, vertex.size,
    FUN = function(x, y, bg, size) {
      # Rotate the square 45 degrees to make it a diamond
      symbols(
        x = x, y = y, bg = bg,
        squares = size,  # Use squares to draw the base shape
        add = TRUE, inches = FALSE,
        fg = bg  # Fill the diamond shape
      )
      # Rotate by 45 degrees
      polygon(c(x, x + size, x, x - size), c(y + size, y, y - size, y),
              col = bg, border = bg)
    }
  )
}

# No clipping, edges will be below the vertices anyway
add_shape("diamond",
  clip = shape_noclip,
  plot = mystar
)

```

Descriptive Summary of Node Attributes and Initial Plot
```{r}
summary(vertex_data)
initialP <- plot(igang, #visualize the network
                 layout = layout_with_fr, #set layout using FR algorithm
                 edge.width = 1,  # Edge width corresponds to relationship level
                 edge.color = "black",  # Color edges by relationship
                 vertex.shape= V(igang)$bshape, #apply color levels for birthplace attribute
                 vertex.color= V(igang)$bcolor, #apply shape levels for birthplace attribute
                 vertex.label = "",  # Show node labels
                 vertex.size = 9,  # Adjust node size
                 main = "Fig. 1: Co-Offending Network of Youth Gang",
                 sub="Data Source: Grund, T. and Densley, J. (2015) Ethnic Homophily and Triad Closure: 
                 Mapping Internal Gang Structure Using Exponential Random Graph Models. 
                 Journal of Contemporary Criminal Justice, Vol. 31, Issue 3, pp. 354-370 "
                 )
legend("bottomleft",                   # Position of the legend (can be adjusted)
       legend = names(birthplace_shapes),  # Names of the Birthplace categories
       pch = c(21, 22, 24, 23),      # Corresponding point shapes for the legend
       pt.cex = 2,                   # Size of the shapes in the legend
       col = birthplace_colors,      # Colors matching the Birthplace categories
       pt.bg = birthplace_colors,    # Background colors of the shapes
       title = "Birthplace",         # Title of the legend
       bty = "n")                    # No box around the legend
```
Attributes Distribution
```{r fig.height=5, fig.width=13}
#Age
p1<- ggplot(vertex_data, aes(x = age)) +
  geom_histogram(aes(y = after_stat(density)), bins = 30, fill = "skyblue2", color = "black", alpha = 0.7) +
  geom_density(fill = "skyblue2", alpha = 0.5) +
  labs(title = "",
       x = "Age",
       y = "Density") +
  theme_minimal()
#Birthplace
vertex_data <- vertex_data %>% #applying country names
  mutate(birthplace = case_when(
    birthplace == 1 ~ "West Africa",
    birthplace == 2 ~ "Caribbean",
    birthplace == 3 ~ "UK",
    birthplace == 4 ~ "East Africa",
    TRUE ~ as.character(birthplace)  # Ensure no NAs if other values exist
  ))
p2<- ggplot(data = vertex_data, aes(x = factor(birthplace))) +
    geom_bar()+
  labs(title = "",
       x = "Birthplace",
       y = "Density") +
  theme_minimal()
#Residence
p3<- ggplot(data = vertex_data, aes(x = factor(residence))) +
    geom_bar()+
  labs(title = "",
       x = "Residence",
       y = "Density") +
  theme_minimal()
#Arrests
p4<- ggplot(vertex_data, aes(x = arrests)) +
   geom_histogram(aes(y = after_stat(density)), bins = 30, fill = "forestgreen", color = "black", alpha = 0.7)+
  geom_density(fill = "forestgreen", alpha = 0.5) +
  labs(title = "",
       x = "Arrests",
       y = "Density") +
  theme_minimal()
#Convictions
p5<- ggplot(vertex_data, aes(x = convictions)) +
   geom_histogram(aes(y = after_stat(density)), bins = 30, fill = "forestgreen", color = "black", alpha = 0.7)+
  geom_density(fill = "forestgreen", alpha = 0.5) +
  labs(title = "",
       x = "Convictions",
       y = "Density") +
  theme_minimal()
#Prison
p6<- ggplot(data = vertex_data, aes(x = factor(prison))) +
    geom_bar()+
  labs(title = "",
       x = "Prison (1=Yes)",
       y = "Density") +
  theme_minimal()

#Music
p7<- ggplot(data = vertex_data, aes(x = factor(music))) +
    geom_bar()+
  labs(title = "",
       x = "Music",
       y = "Density") +
  theme_minimal()
#Ranking
p8<- ggplot(vertex_data, aes(x = rank)) +
   geom_histogram(aes(y = after_stat(density)), bins = 30, fill = "red", color = "black", alpha = 0.7)+
  geom_density(fill = "red", alpha = 0.5) +
  labs(title = "",
       x = "Ranking",
       y = "Density") +
  theme_minimal()
annotate_figure(ggarrange(p1,p2,p3,p4, p5, p6, p7, p8,
                          ncol = 4, nrow = 2, 
                          labels = c("Age", "Birthplace","Residence","Arrests", "Convictions", "Prison", "Music", "Rank")))
```
Relationship Plot
```{r}
plot(igang, #visualize the network
     layout = layout_with_fr, #set layout using FR algorithm
     edge.width = 1,  # Edge width corresponds to relationship level
     edge.color = edge_colors[E(igang)$relationship],  # Color edges by relationship
     vertex.label = "",  # Show node labels
     vertex.size = 7,  # Adjust node size
     vertex.shape= V(igang)$bshape, #apply color levels for birthplace attribute
     vertex.color= V(igang)$bcolor, #apply shape levels for birthplace attribute
     main = "Fig. 2: London Street Gang Network with Relationship Levels",
     sub="co-offend together= Blue
     co-offend together, serious crime= Yellow
     co-offend together, serious crime, kin= Red"
     )
legend("bottomleft",                   # Position of the legend (can be adjusted)
       legend = names(birthplace_shapes),  # Names of the Birthplace categories
       pch = c(21, 22, 24, 23),      # Corresponding point shapes for the legend
       pt.cex = 2,                   # Size of the shapes in the legend
       col = birthplace_colors,      # Colors matching the Birthplace categories
       pt.bg = birthplace_colors,    # Background colors of the shapes
       title = "Birthplace",         # Title of the legend
       bty = "n")                    # No box around the legend
```

*Replication*
```{r}
library(statnet) #load statnet
library(intergraph)
```
Network from Network Package
```{r}
gang <- as.network(as.matrix(adj>1),directed=F) #create network object
network::set.vertex.attribute(gang,names(att),att) # Add the vertex attributes into the network
```

```{r}
plot(gang,
     displaylabels=T,
     label.cex=.5,
     edge.col=rgb(150,150,150,100,maxColorValue=255),
     displayisolates=F)               
```


ERGM: Homophily
Model 1
```{r}
model1 <- ergm(gang ~ edges + nodematch("Birthplace")) #Model 1 replication
model2 <- ergm(gang ~ edges + nodematch("Birthplace") + gwesp(0, fixed = TRUE)) #Model 2 Replication
model4 <- ergm(gang ~ edges + nodematch("Birthplace") + nodefactor("Birthplace") + gwesp(0, fixed = TRUE)) #Model 2 Replication
```

All Model Summary and Goodness of Fit
```{r}
#All Models
screenreg(list(model1, model2, model3), digits = 3, caption = "ERGM", label = "ergm",
       custom.coef.map = list("edges" = "Edges", 
                              "nodematch.Birthplace" = "Match(Birthplace)",
                              "nodefactor.Birthplace.2"="Caribean",
                              "nodefactor.Birthplace.3"="UK",
                              "nodefactor.Birthplace.4"="East Africa",
                              "gwesp.fixed.0"= "GWESP"
       ))
#Goodness of Fit
#Assess model fit-m2
gof<- ergm::gof(model3 ~ model + deg + esp + triadcensus + distance)
#visualize goodness-of-fit measures 
ergm:::plot.gof(gof)
```

